Approximation as Differentiation

### 2025-07-02 main/approx-as-tangents.tex
Rough pass over Stable Domain Theory.
JAX citation.

Derivatives can be computed numerically by computing $f$ on small perturbations of its input, or symbolically by examining a closed-form representation of $f$. However, a more common and practical technique is to use \emph{automatic differentiation}, where a program computing $f$ is instrumented to produce (a representation of) the forward and/or backward derivative as a side-effect of producing the output. \bob{cite something on autodiff}. This has led to the area of differentiable programming, where programming languages and their implementations are specifically designed to admit efficient automatic differentation algorithms. \bob{cite: JAX, TensorFlow, Dex, Jesse Sigal, etc.}

@@

Stability is a property that can be required of monotone functions that was invented by Berry \cite{berry86} in an attempt to capture sequentiality. This was not successful (see the $\mathrm{gustave}$ function in \exref{parallel-or}), but we will see now how it is closely related to problem of computing forward and backwards maps of approximations needed in \GPS. A textbook description of stable functions in the context of domain theory is given by \citet{amadio-curien} (Chapter 12). We start with a property that is weaker than stability, but is easier to motivate in connection with derivatives of smooth functions.

>>>>

Derivatives can be computed numerically by computing $f$ on small perturbations of its input, or symbolically by examining a closed-form representation of $f$. However, a more common and practical technique is to use \emph{automatic differentiation}, where a program computing $f$ is instrumented to produce (a representation of) the forward and/or backward derivative as a side-effect of producing the output. \bob{cite something on autodiff}. This has led to the area of differentiable programming, where programming languages and their implementations are specifically designed to admit efficient automatic differentation algorithms~\cite{jax2018github}. \bob{cite: TensorFlow, Dex, Jesse Sigal, etc.}

@@

Stability is a property that can be required of monotone functions that was invented by \citet{berry79} in an attempt to capture sequentiality. This was not successful (see the $\mathrm{gustave}$ function in \exref{parallel-or}), but we will see now how it is closely related to problem of computing forward and backwards maps of approximations needed in \GPS. A textbook description of stable functions in the context of domain theory is given by \citet{amadio-curien} (Chapter 12). We start with a property that is weaker than stability, but is easier to motivate in connection with derivatives of smooth functions.

### 2025-07-02 main/introduction.tex
JAX citation.

Perera and collaborators introduced the idea of {\em Galois program slicing} as a particular conception of program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}. Galois program slicing (hereafter simply {\emph \GPS}) forms the basis of the open source data visualisation tool \href{https://f.luid.org/}{Fluid} that allows interactive exploration of programmatically generated visualisations.

>>>>

Perera and collaborators introduced the idea of {\em Galois program slicing} as a particular conception of program slicing for provenance, described in several publications~\cite{perera12a,perera16d,ricciotti17}. Galois program slicing (hereafter simply {\emph \GPS}) forms the basis of the open source data visualisation tool \href{https://f.luid.org/}{Fluid}~\cite{perera2025fluid} that allows interactive exploration of programmatically generated visualisations.
